Beam-Helicity and Beam-Charge Asymmetries \\ in the Bethe-Heitler and DVCS processes on \\ an Unpolarised Proton Target \underline{Paper Tag: PDVCS0607 (v1.3)} \underline{Drafting Committee: DC90 (\today)}

### 2011-09-07 background_and_systematic_dual.tex

The total systematic uncertainty for the 2006-2007 data sample was
determined for each kinematic bin by adding in quadrature the
  uncertainties arising from the background correction, the
  missing-mass shifts, and the detector effects.  The 1996-2005 sample also has a
systematic uncertainty from misalignment of the
spectrometer~\cite{Air09}, which has been eliminated for the 2006-2007 data sample due to improved
  surveying measurements~\cite{Bur10}. The contribution from
each uncertainty to each amplitude extracted in a single, overall kinematic bin
from the 2006-2007 data  is summarised in Table
\ref{table_systematic_contributions_0607}.

@@

The data sample in the exclusive region contains events involving both
production of real photons in which the proton remains intact and
events involving the excitation of the target proton to a $\Delta^+$
resonant state (``associated production''). The recoiling proton is not detected and
  the calorimeter resolution does not allow separation of the latter events from the rest of the data sample.
No systematic uncertainty is assigned for the contribution from these
events; they are treated as part of the signal. A Monte Carlo
calculation based on the parameterisation from ref.~\cite{Bra76} is
used to estimate the contribution to the event sample from resonant
production in each kinematic bin. The results, called the associated fractions and labelled ``Assoc. fraction'', are shown in the last row of each figure in the results section. The method used to perform this estimation is described in detail in
ref.~\cite{Air08}.

>>>>

The total systematic uncertainty for the 2006-2007 data sample was
determined for each kinematic bin by adding in quadrature the
  uncertainties arising from the background correction, the
  missing-mass shifts, and the detector effects.  The 1996-2005 sample also has a
systematic uncertainty from misalignment of the
spectrometer~\cite{Air09}, which has been eliminated for the 2006-2007 data sample due to improved
  surveying measurements~\cite{Bur10}. The contribution from
each uncertainty to each amplitude extracted in a single, overall kinematic bin
from the 2006-2007 data  is summarised in Table~\ref{table_systematic_contributions_0607}.

@@

The data sample in the exclusive region contains events involving both
production of real photons in which the proton remains intact and
events involving the excitation of the target proton to a $\Delta^+$
resonant state (``associated production''). The recoiling proton is not detected and
  the calorimeter resolution does not allow separation of the latter events from the rest of the data sample.
No systematic uncertainty is assigned for the contribution from these
events; they are treated as part of the signal. A Monte Carlo
calculation based on the parameterisation from ref.~\cite{Bra76} is
used to estimate the \red{fractional} contribution to the event sample from resonant
production in each kinematic bin. The results, called the associated fractions and labelled ``Assoc. fraction'', are shown in the last row of figure\red{s} in the results section. The method used to perform this estimation is described in detail in
ref.~\cite{Air08}.

### 2011-09-07 experiment_and_data.tex

The asymmetry amplitudes $A_{\textrm{LU,I/DVCS}}^{\sin(n\phi)}$ and
$A_{\textrm{C}}^{\cos(n\phi)}$ relate respectively to the Fourier
coefficients $s_{n,\textrm{unp}}^{W}$ and $c_{n,\textrm{unp}}^{I}$ from the interference and
DVCS terms in Eqs.~\ref{e:alui}-\ref{e:ac}. The asymmetry amplitudes
may also be affected by the lepton propagators and the other
$\phi$-dependent terms in the denominators in
Eqs.~\ref{e:alui}-\ref{e:ac}. The  $s_{n,\textrm{unp}}^{W}$ and $c_{n,\textrm{unp}}^{V}$ Fourier
coefficients depend on ``$\mathcal{C}$-functions'' \cite{{Bel02b}}, each of which is a
combination of complex Compton Form Factors (CFFs). A CFF is a
convolution of a GPD with a hard scattering kernel. Contributions of
GPDs to the cross section may be relatively suppressed by powers of
$1/Q$, the origin of which may be kinematic in nature or due to the
twist level of the GPD. Leading twist is twist-2. Typically, the
contribution of a twist-$n$ GPD, and hence the corresponding CFF, is
suppressed by $\mathcal{O}(1/Q^{n-2})$. Table~\ref{tab_amplitudes}
presents the asymmetry amplitudes extracted in this analysis and the
related Fourier coefficients, dominant $\mathcal{C}$-function, twist-level and level of suppression.

At H{\sc ermes} kinematic conditions, the leading-twist beam-helicity asymmetry amplitude is $A_{\textrm{LU,I}}^{\sin\phi}$ and the leading beam-charge asymmetry amplitude is $A^{\cos\phi}_{\textrm{C}}$. These asymmetry amplitudes, for an unpolarised target, are sensitive respectively to Fourier coefficients $s_{1,\textrm{unp}}^{\textrm{I}}$ and $c_{1,\textrm{unp}}^{\textrm{I}}$. Both of these Fourier coefficients have a dominant contribution from the $\mathcal{C}_{\textrm{unp}}^{\textrm{I}}$-function:
\begin{eqnarray}
s_{1,\textrm{unp}}^{\textrm{I}} \approx&\,\,\,\,8\lambda ky(2-y)\mathfrak{Im}\mathcal{C}_{\textrm{unp}}^{\textrm{I}}\quad\textrm{and}&\label{eq:s1}
\\
 c_{1,\textrm{unp}}^{\textrm{I}} \approx&8k(2- 2y + y^{2})\mathfrak{Re}\mathcal{C}_{\textrm{unp}}^{\textrm{I}}\,,&\label{eq:c1}
\end{eqnarray}
where the definition of the kinematic factor $k$ is found
  in ref.\cite{Bel02b}. The $\mathcal{C}_{\textrm{unp}}^{\textrm{I}}$-function can be
written
\cite{Bel02b} 
\begin{equation}
 \mathcal{C}_{\textrm{unp}}^{\textrm{I}} = F_{1}\mathcal{H} + \frac{x_{\textrm{B}}}{2-x_{\textrm{B}}}(F_{1}+F_{2})\widetilde{\mathcal{H}} -\frac{t}{4M^{2}}F_{2}\mathcal{E},
\label{Eq_cunp}
\end{equation}
where $F_{1}$ and $F_{2}$ are respectively the Dirac and Pauli form
factors of the nucleon and $\mathcal{H}$, $\widetilde{\mathcal{H}}$ and
$\mathcal{E}$ are CFFs that relate respectively to the GPDs $H$,
$\widetilde{H}$ and $E$.  At H{\sc ermes} kinematic
conditions (where $x_{\textrm{B}}$ and $\frac{-t}{4M^2}$ are of order 0.1), the
contributions of CFFs $\widetilde{\mathcal{H}}$ and $\mathcal{E}$ can be
neglected in Eq.~\ref{Eq_cunp} (in first approximation) with respect to $\mathcal{H}$ since they
are kinematically suppressed by an order of magnitude or more.
Hence, the behaviour of
$\mathcal{C}_{\textrm{unp}}^{\textrm{I}}$ is determined by CFF $\mathcal{H}$
and therefore GPD $H$ can be constrained through
measurements of $A_{\textrm{LU,I}}^{\sin\phi}$ and $A^{\cos\phi}_{\textrm{C}}$.
The asymmetry amplitudes $A_{\textrm{LU},\textrm{I}}^{\sin(2\phi)}$,
$A^{\cos(0\phi)}_{\textrm{C}}$ and $A^{\cos(2\phi)}_{\textrm{C}}$ also depend
upon this $\mathcal{C}$-function. However, the $A^{\cos(0\phi)}_{\textrm{C}}$ amplitude is kinematically suppressed compared to the $A^{\cos\phi}_{\textrm{C}}$ amplitude, and the  $A_{\textrm{LU,I}}^{\sin(2\phi)}$ and $A^{\cos(2\phi)}_{\textrm{C}}$ amplitudes are higher-twist quantities (see Table~\ref{tab_amplitudes}).

>>>>

The asymmetry amplitudes $A_{\textrm{LU,I/DVCS}}^{\sin(n\phi)}$ and
$A_{\textrm{C}}^{\cos(n\phi)}$ relate respectively to the Fourier
coefficients $s_{n,\textrm{unp}}^{W}$ and $c_{n,\textrm{unp}}^{I}$ from the interference and
DVCS terms in eqs.~\ref{e:alui}-\ref{e:ac}. The asymmetry amplitudes
may also be affected by the lepton propagators and the other
$\phi$-dependent terms in the denominators in
eqs.~\ref{e:alui}-\ref{e:ac}. The  $s_{n,\textrm{unp}}^{W}$ and $c_{n,\textrm{unp}}^{V}$ Fourier
coefficients depend on ``$\mathcal{C}$-functions'' \cite{{Bel02b}}, each of which is a
combination of complex Compton Form Factors (CFFs). A CFF is a
convolution of a GPD with a hard scattering kernel. Contributions of
GPDs to the cross section may be relatively suppressed by powers of
$1/Q$, the origin of which may be kinematic in nature or due to the
twist level of the GPD. Leading twist is twist-2. Typically, the
contribution of a twist-$n$ GPD, and hence the corresponding CFF, is
suppressed by $\mathcal{O}(1/Q^{n-2})$. Table~\ref{tab_amplitudes}
presents the asymmetry amplitudes extracted in this analysis and the
related Fourier coefficients, dominant $\mathcal{C}$-function, twist-level and level of suppression.

At H{\sc ermes} kinematic conditions, the leading-twist beam-helicity asymmetry amplitude is $A_{\textrm{LU,I}}^{\sin\phi}$ and the leading beam-charge asymmetry amplitude is $A^{\cos\phi}_{\textrm{C}}$. These asymmetry amplitudes, for an unpolarised target, are sensitive respectively to Fourier coefficients $s_{1,\textrm{unp}}^{\textrm{I}}$ and $c_{1,\textrm{unp}}^{\textrm{I}}$. Both of these Fourier coefficients have a dominant contribution from the $\mathcal{C}_{\textrm{unp}}^{\textrm{I}}$-function:
\begin{eqnarray}
s_{1,\textrm{unp}}^{\textrm{I}} \approx&\,\,\,\,8\lambda ky(2-y)\mathfrak{Im}\mathcal{C}_{\textrm{unp}}^{\textrm{I}}\quad\textrm{and}&\label{eq:s1}
\\
 c_{1,\textrm{unp}}^{\textrm{I}} \approx&8k(2- 2y + y^{2})\mathfrak{Re}\mathcal{C}_{\textrm{unp}}^{\textrm{I}}\,,&\label{eq:c1}
\end{eqnarray}
where the definition of the kinematic factor $k$ is found
  in ref.\cite{Bel02b}. The $\mathcal{C}_{\textrm{unp}}^{\textrm{I}}$-function can be
written
\cite{Bel02b} 
\begin{equation}
 \mathcal{C}_{\textrm{unp}}^{\textrm{I}} = F_{1}\mathcal{H} + \frac{x_{\textrm{B}}}{2-x_{\textrm{B}}}(F_{1}+F_{2})\widetilde{\mathcal{H}} -\frac{t}{4M^{2}}F_{2}\mathcal{E},
\label{Eq_cunp}
\end{equation}
where $F_{1}$ and $F_{2}$ are respectively the Dirac and Pauli form
factors of the nucleon and $\mathcal{H}$, $\widetilde{\mathcal{H}}$ and
$\mathcal{E}$ are CFFs that relate respectively to the GPDs $H$,
$\widetilde{H}$ and $E$.  At H{\sc ermes} kinematic
conditions (where $x_{\textrm{B}}$ and $\frac{-t}{4M^2}$ are of order 0.1), the
contributions of CFFs $\widetilde{\mathcal{H}}$ and $\mathcal{E}$ can be
neglected in eq.~\ref{Eq_cunp} (in first approximation) with respect to $\mathcal{H}$ since they
are kinematically suppressed by an order of magnitude or more.
Hence, the behaviour of
$\mathcal{C}_{\textrm{unp}}^{\textrm{I}}$ is determined by CFF $\mathcal{H}$
and therefore GPD $H$ can be constrained through
measurements of $A_{\textrm{LU,I}}^{\sin\phi}$ and $A^{\cos\phi}_{\textrm{C}}$.
The asymmetry amplitudes $A_{\textrm{LU},\textrm{I}}^{\sin(2\phi)}$,
$A^{\cos(0\phi)}_{\textrm{C}}$ and $A^{\cos(2\phi)}_{\textrm{C}}$ also depend
upon this $\mathcal{C}$-function. However, the $A^{\cos(0\phi)}_{\textrm{C}}$ amplitude is kinematically suppressed compared to the $A^{\cos\phi}_{\textrm{C}}$ amplitude, and the  $A_{\textrm{LU,I}}^{\sin(2\phi)}$ and $A^{\cos(2\phi)}_{\textrm{C}}$ amplitudes are higher-twist quantities (see Table~\ref{tab_amplitudes}).

### 2011-09-07 results_dual.tex

The results of the beam-helicity and beam-charge asymmetry amplitudes extracted from the complete 1996-2007 hydrogen sample are shown in figures~\ref{bsa_xbjrange} and \ref{bca_xbjrange}. The number of analysable events available from the 2006-2007 data set (70352) is approximately three times greater than the number of events recorded in the 1996-2005 sample (24817). The  asymmetry amplitudes extracted from the complete 1996-2007 data set thus resemble the 2006-2007 result. This resemblance is not so evident for the beam-helicity asymmetry amplitudes because the beam polarisation was lower in 2006 and 2007 than in 1996-2005, giving the 2006\red{-}2007 data a lower weighting in the combined fit. \red{The systematic error bands associated with the asymmetry amplitudes extracted from the combined data set were determined using Monte Carlo simulations as explained in section~4.}

@@

The $A_{\textrm{C}}^{\cos(n\phi)}$ amplitudes are shown in figure~\ref{bca_xbjrange}. The systematic uncertainties are estimated in the same procedure as \red{used to estimate the systematic uncertainties of} the beam-helicity asymmetries. The leading-twist $A_{\textrm{C}}^{\cos(0\phi)}$ and $A_{\textrm{C}}^{\cos\phi}$ amplitudes are both non-zero. There is an expected relationship between $A_{\textrm{C}}^{\cos(0\phi)}$ and $A_{\textrm{C}}^{\cos\phi}$ as they \red{depend on the same $\mathcal{C}$-function. The kinematic suppression of $A_{\textrm{C}}^{\cos(0\phi)}$ with regards to $A_{\textrm{C}}^{\cos\phi}$ can be found in table~\ref{tab_amplitudes}. The measured} amplitudes are found to diverge with opposite sign from zero at increasing values of $-t$ but they have no discernible dependence on $x_{\textrm{B}}$ and $Q^{2}$. The $A_{\textrm{C}}^{\cos(2\phi)}$ and $A_{\textrm{C}}^{\cos(3\phi)}$ amplitudes are both consistent with zero and have no significant variation in value over the range in $-t$, $x_{\textrm{B}}$ and $Q^{2}$. The $A_{C}^{\cos(2\phi)}$ amplitude is related to twist-3 GPDs and $A_{\textrm{C}}^{\cos(3\phi)}$ relates to gluon helicity-flip GPDs. Both of these amplitudes are expected to be suppressed at H{\sc ermes} kinematic conditions compared to the leading-twist amplitudes.

@@

In order to provide more detailed information that can be used in future fits, in particular for the determination of the entanglement of the skewness and $-t$ dependences of GPDs, the amplitudes already presented in figures~\ref{bsa_xbjrange} and~\ref{bca_xbjrange} are shown as a function of $-t$ for three different ranges of $x_{\textrm{B}}$ in figures~\ref{bsa_xbjrange2} and~\ref{bca_xbjrange2}. \red{These figures represent the kinematic dependences of the amplitudes in a less-correlated manner than the one dimensional projections}. There is no observed $-t$ dependence of any of the $\mathcal{A}_{\textrm{LU}}$ amplitudes in any of the distinct ranges of $x_{\textrm{B}}$. No additional features or $-t$ dependences are observed for the $\mathcal{A}_{\textrm{C}}$ amplitudes in these separate $x_{\textrm{B}}$ ranges.

>>>>

The results of the beam-helicity and beam-charge asymmetry amplitudes extracted from the complete 1996-2007 hydrogen sample are shown in Figs.~\ref{bsa_xbjrange} and \ref{bca_xbjrange}. The number of analysable events available from the 2006-2007 data set (70352) is approximately three times greater than the number of events recorded in the 1996-2005 sample (24817). The  asymmetry amplitudes extracted from the complete 1996-2007 data set thus resemble the 2006-2007 result. This resemblance is not so evident for the beam-helicity asymmetry amplitudes because the beam polarisation was lower in 2006 and 2007 than in 1996-2005, giving the 2006-2007 data a lower weighting in the combined fit. The systematic error bands associated with the asymmetry amplitudes extracted from the combined data set were determined using Monte Carlo simulations as explained in section~4. The missing-mass uncertainty for the combined data set was calculated using the procedure described in that section and the results were added in quadrature.

@@

The $A_{\textrm{C}}^{\cos(n\phi)}$ amplitudes are shown in figure~\ref{bca_xbjrange}. The systematic uncertainties are estimated in the same procedure as \red{used to estimate the systematic uncertainties of} the beam-helicity asymmetries. The leading-twist $A_{\textrm{C}}^{\cos(0\phi)}$ and $A_{\textrm{C}}^{\cos\phi}$ amplitudes are both non-zero. There is an expected relationship between $A_{\textrm{C}}^{\cos(0\phi)}$ and $A_{\textrm{C}}^{\cos\phi}$ as they depend on the same $\mathcal{C}$-function. The kinematic suppression of $A_{\textrm{C}}^{\cos(0\phi)}$ with regards to $A_{\textrm{C}}^{\cos\phi}$ can be found in table~\ref{tab_amplitudes}. The measured amplitudes are found to diverge with opposite sign from zero at increasing values of $-t$ but they have no discernible dependence on $x_{\textrm{B}}$ and $Q^{2}$. The $A_{\textrm{C}}^{\cos(2\phi)}$ and $A_{\textrm{C}}^{\cos(3\phi)}$ amplitudes are both consistent with zero and have no significant variation in value over the range in $-t$, $x_{\textrm{B}}$ and $Q^{2}$. The $A_{C}^{\cos(2\phi)}$ amplitude is related to twist-3 GPDs and $A_{\textrm{C}}^{\cos(3\phi)}$ relates to gluon helicity-flip GPDs. Both of these amplitudes are expected to be suppressed at H{\sc ermes} kinematic conditions compared to the leading-twist amplitudes.

@@

In order to provide more detailed information that can be used in future fits, in particular for the determination of the entanglement of the skewness and $-t$ dependences of GPDs, the amplitudes already presented in Figs.~\ref{bsa_xbjrange} and~\ref{bca_xbjrange} are shown as a function of $-t$ for three different ranges of $x_{\textrm{B}}$ in Figs.~\ref{bsa_xbjrange2} and~\ref{bca_xbjrange2}. These figures represent the kinematic dependences of the amplitudes in a less-correlated manner than the one dimensional projections. There is no observed $-t$ dependence of any of the $\mathcal{A}_{\textrm{LU}}$ amplitudes in any of the distinct ranges of $x_{\textrm{B}}$. No additional features or $-t$ dependences are observed for the $\mathcal{A}_{\textrm{C}}$ amplitudes in these separate $x_{\textrm{B}}$ ranges.

### 2011-09-07 theory.tex

\section{Introduction}
Generalised Parton Distributions (GPDs)
\cite{Mue94,Ji97a,Rad97} encompass the familiar parton
distribution functions (PDFs) and nucleon form factors to provide a
comprehensive description of the structure of the nucleon.
Although GPDs are difficult to access experimentally, the H{\sc ermes}
collaboration has previously published results
\cite{Air01,Air06,Air08,Air09,Air10, Air10a, Air10b, Air11} on
the leptoproduction of real photons from a nucleon or nucleus that can
be used to constrain GPDs via Compton Form Factors (CFFs). A thorough
description of the nucleon in terms of GPDs would allow the deduction
of the total angular momentum of partons in the nucleon, and the
construction of a longitudinal-momentum-dissected transverse spatial map of parton positions.

@@

Exclusive leptoproduction of real photons
($e\,p\,\rightarrow\,e'\,p'\,\gamma$) arises from
two experimentally indistinguishable processes shown in
Fig.~\ref{spin}: the Deeply Virtual Compton Scattering (DVCS) process,
which is the emission of a real photon by the struck quark from the nucleon, and the Bethe-Heitler (BH) process, which is elastic lepton-proton scattering with the
emission of a Bremsstrahlung photon by the lepton. 
The BH process is calculable in the QED framework; this process is
dominant at the kinematic conditions of the H{\sc ermes} experiment, but the
scattering amplitudes of the two processes interfere and the large BH amplitude
amplifies the contribution of the DVCS amplitude to the interference term. 
It is through the study of this interference term at H{\sc ermes} that
useful information for the constraint of certain GPDs can be obtained \cite{Bel02b}.

>>>>

\section{Introduction}
Generalised Parton Distributions (GPDs)
\cite{Mue94,Ji97a,Rad97} encompass the familiar Parton
Distribution Functions (PDFs) and nucleon Form Factors (FFs) to provide a
comprehensive description of the structure of the nucleon.
Although GPDs are difficult to access experimentally, the H{\sc ermes}
collaboration has previously published results
\cite{Air01,Air06,Air08,Air09,Air10, Air10a, Air10b, Air11} on
the leptoproduction of real photons from a nucleon or nucleus that can
be used to constrain GPDs via Compton Form Factors (CFFs). A thorough
description of the nucleon in terms of GPDs would allow the deduction
of the total angular momentum of partons in the nucleon, and the
construction of a longitudinal-momentum-dissected transverse spatial map of parton positions.

@@

Exclusive leptoproduction of real photons
($e\,p\,\rightarrow\,e'\,p'\,\gamma$) arises from
two experimentally indistinguishable processes shown in
figure~\ref{spin}: the Deeply Virtual Compton Scattering (DVCS) process,
which is the emission of a real photon by the struck quark from the nucleon, and the Bethe-Heitler (BH) process, which is elastic lepton-proton scattering with the
emission of a Bremsstrahlung photon by the lepton. 
The BH process is calculable in the QED framework; this process is
dominant at the kinematic conditions of the H{\sc ermes} experiment, but the
scattering amplitudes of the two processes interfere and the large BH amplitude
amplifies the contribution of the DVCS amplitude to the interference term. 
It is through the study of this interference term at H{\sc ermes} that
useful information for the constraint of certain GPDs can be obtained \cite{Bel02b}.
